---
title: "Coding workshop: Week 10"
author: "Sara Gonzalez"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
execute:
  message: false
  warning: false
---

```{r set-up}
# model packages
library(MASS) # have to read this in before tidyverse 
library(lme4)
library(glmmTMB) # ok if you don't have this - just comment it out

# diagnostics and model info
library(DHARMa)
library(MuMIn)
library(ggeffects)
library(lmtest) 
library(broom)

# general usage
library(tidyverse)
library(here)
library(naniar)
library(skimr)
library(GGally)
library(flextable)

salamanders <- read_csv(here("data", "Salamanders.csv"))
```

# Data info from `glmmTMB`:

`site`: name of a location where repeated samples were taken  
`mined`: factor indicating whether the site was affected by mountain top removal coal mining  
`cover`: amount of cover objects in the stream (scaled)  
`sample`: repeated sample  
`DOP`: Days since precipitation (scaled)  
`Wtemp`: water temperature (scaled)  
`DOY`: day of year (scaled)  
`spp`: abbreviated species name, possibly also life stage  
`count`: number of salamanders observed  

scaling is useful when its very different numbers (ex: 15-17 Celsius vs 1-365 days)
takes mean of continuous variables and shifts everything else to center on mean (variables are on same scale so easier to interpret). neg # = below mean and pos # = above mean. 

# Explore the data set:

## histogram of counts:

```{r histogram}
ggplot(data = salamanders, aes(x = count)) +
  geom_histogram(bins = 17)
```

## Missingness:

```{r missingness}
gg_miss_var(salamanders)
```

## Skim: 

```{r skim-dataset}
skim(salamanders)
```

## Pairs plot:  

```{r pairs-plot}
#| fig.width: 12
#| fig.height: 12
#| out.width: 90%
#| fig.align: center

salamanders %>% 
  # take out site column
  select(!site) %>% 
  ggpairs()

# water temp might be normally distributed in relation to species; DOY + sample pearson's correlation is high so indicates strong relationship, but meaningfully, sample isn't necessarily measured variable (more for identification); DOP + wtemp -> negative relationship that flattens out, weak correlation but relationship exists
```

# Question: How does salamander count vary with mined status, species, and stream cover?
does mining affect # salamanders
## Build models

```{r models}
# linear model, we know this is wrong, not continuous variable (discrete)
salmod1 <- lm(count ~ cover + mined + spp, data = salamanders)

# generalized linear model with Poisson distribution; assumption : variance = mean
salmod2 <- glm(count ~ cover + mined + spp, data = salamanders, family = "poisson")

# generalized linear model with negative binomial distribution; glm.nb assumes nb dist.; appropriate for over dispersion, variance > mean
salmod3 <- glm.nb(count ~ cover + mined + spp, data = salamanders)

# generalized linear model with Poisson distribution and random effect of site


# generalized linear model with negative binomial distribution and random effect of site

```

## Look at residuals

```{r model-residuals}
#| fig.height: 6
#| fig.width: 8
#| fig.align: center

# check diagnostics, use darma b/c GLM don't follow same assumptions, simulation of residuals
simulateResiduals(salmod1, plot = TRUE)
simulateResiduals(salmod2, plot = TRUE) # look better but significant deviation, residuals are over-dispersed
simulateResiduals(salmod3, plot = TRUE)

```

## Which distribution to use?

```{r choosing-model}
MuMIn::model.sel(salmod1, salmod2, salmod3) # 100% chance of best model = 1 (weight)
```

## Model summary

```{r summaries}
# model object
salmod3

# summary (z value = test stat based on z dist.)
summary(salmod3)
  # cover is potentially not significant, sppDM, sppEC-L as well keeping everything else constant
# confidence intervals
confint(salmod3)

# adjusted R2
r.squaredGLMM(salmod3) # only look at delta, amount of variance = 43%
# fairly low due to 3 different predictors
```

```{r}
# model object in table
salmod3 %>% 
  as_flextable()

tidy(salmod3) %>% 
  flextable()
```


## Visualizing:

```{r visualization}
#| fig.width: 12
#| fig.height: 10
#| out.width: 90%
#| fig.align: center

plot(ggpredict(salmod3, terms = c("cover", "mined")), add.data = TRUE)

plot(ggpredict(salmod3, terms = c("cover", "mined", "spp")), add.data = TRUE)
# cover has no significant effect on salamander count, but mined status and species do
```

